Dish: Distribution-oblivious Shell Scripting

### 2019-11-23 pldi/draft.tex
Almost done with section 5

In order to expose data parallelism from the output of a command,
files that don't point to a resource can have an upper bound on the
lines that they can tranfer. The following example illustrates why
this is essential.

\TODO{Add an example of a command that has two outputs, one of which
  goes to one command, while another goes to a different
  command. Maybe only show the graph here and not the command itself.}

@@

% \begin{figure}[t]
% \centering
\includegraphics[width=0.25\textwidth]{\detokenize{./figs/dish_g1.pdf}}

\includegraphics[width=0.15\textwidth]{\detokenize{./figs/dish_g2.pdf}}
% \caption{
%   \textbf{\sys transformation overview.}
%   Given a shell pipeline, \sys leverages command distributability to identify
%   high-distributability stages, rewrite them, and orchestrate their execution.
% }
% \vspace{-15pt}
% \label{fig:overview}
% \end{figure}

@@

The \ttt{Dish} optimizer is given a dataflow graph representing a
shell program and outputs an optimized dataflow graph, where data
parallelism opportunities have been exploited. More concretely, the
optimizer starts by splitting input files in several chunks, given the
desirable output dataflow graph width. It then starts from the source
nodes of the graph, iteratively performing the graph transformations
that were described in \Cref{ir:transformations}. If the incoming
edges of the node are less than the desirable graph width, it performs
edge splits, and then it performs maximal node splits. The final graph
can then be implemented by spawning a process for each node, and
redirecting the outputs according to the graph edges.

\nv{Give example, extend}

\kk{Mention sort, wc}

@@

It is common in the dataflow graph for a node to have many incoming
edges, or many outgoing edges (when all but the last outgoing edges
are bounded). However some nodes, might only read their input from one
source (e.g. stdin). In this case, the distributed implementation
needs to concatenate input files (or conversely split output
files). Here is an example:

\kk{Show an example graph! of cat | tr | sort. The command in the
  middle must only take input from one file (maybe just stdin).}

We address this issue using primitive shell constructs for file
manipulation. More concretely a merger can be implemented by a simple
\verb|cat|. Given a set of input files \verb|in1, in2, ...| and an
output file \verb|out|, a merger can be implemented as:

>>>>

In order to expose data parallelism from the output of a command,
files that don't point to a resource can have an upper bound on the
lines that they can tranfer. This can be useful to utilize avaliable
parallelism after encountering a command that cannot be parallelized
(e.g. \ttt{sha1sum}). The following example illustrates this scenario.

\includegraphics[width=0.40\columnwidth]{\detokenize{./figs/dish_g2.pdf}}

@@

@@

The \ttt{Dish} optimizer is given a dataflow graph representing a
shell program and outputs an optimized dataflow graph, where data
parallelism opportunities have been exploited. More concretely, the
optimizer starts by splitting input files in several chunks, given the
desirable output dataflow graph width. It then starts from the source
nodes of the graph, iteratively performing the graph transformations
that were described in \Cref{ir:transformations}. If the incoming
edges of the node are less than the desirable graph width, it performs
edge splits, and then it performs maximal node splits. The final graph
can then be implemented by spawning a process for each node, and
redirecting the outputs according to the graph edges. An example
transformation can be seen in \Cref{fig:optimizer-transformation}.

\begin{figure}[t]
\centering
\includegraphics[width=0.70\columnwidth]{\detokenize{./figs/dish_g1.pdf}}
\caption{
  \textbf{Example graph transformation.}
  The optimizer splits a stateless node to utilize
  available data parallelism.
}
\vspace{-15pt}
\label{fig:optimizer-transformation}
\end{figure}

The optimizer also applies custom transformations on several commands
to further improve performance. Two notable examples are the
transformations on \ttt{sort} and \ttt{wc}. Occurences of \ttt{sort}
that have more than one input are tranformed to \ttt{sort} commands on
each input, followed by a \ttt{sort -m}, to produce the final sorted
result. The transformation on occurences of \ttt{wc} that have more
than input \ttt{i1, t2} is illustrated below as a shell script:

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
  paste -d '+' \
   <(wc i1 | tr -s ' ' '\n' | tail -n +2) \
   <(wc i2 | tr -s ' ' '\n' | tail -n +2) |
   bc | tr -s '\n' ' ' |
   sed 's/^/   /' | sed 's/$/\ /'
\end{lstlisting}

The above script takes the outputs produced by two \ttt{wc} commands
and aggregates them accordingly.

@@

It is common in the dataflow graph for a node to have many incoming
edges, or many outgoing edges (when all but the last outgoing edges
are bounded). However some nodes, might only read their input from one
source (e.g. stdin). In this case, the distributed implementation
needs to concatenate input files (or conversely split output
files).
%
%% \tr{Show an example graph! of cat | tr | sort. The command in the
%%   middle must only take input from one file (maybe just stdin).}
%
We address this issue using primitive shell constructs for file
manipulation. More concretely a merger can be implemented by a simple
\verb|cat|. Given a set of input files \verb|in1, in2, ...| and an
output file \verb|out|, a merger can be implemented as:
